How Beauty Trumps Performance: An Examination of Instructor Evaluations at the University of Texas

### 2023-04-21 outputs/paper/paper.qmd
plots with ggplot2

# 2.3 Data Analysis 
The data set from AER was gathered by installing the AER package [@AER] in the RStudio [@rstudio] console, and then the package contents were extracted to only include the TeachingRatings data set in the "01-download_data.R" file under "/scripts". The data was then analyzed using the R programming language [@rcoreteam2022]. Additionally, the programming packages dplyr [@wickham2023] as well as tidyverse [@wickham2019], and [@testthat] were used to analyze the data, test and create graphs.

@@

# males graph of beauty and evaluation 
males <- subset(TeachingRatings, gender == "male") |> select(eval, beauty, gender)
plot(eval ~ beauty, data = males, ylim = c(2.0, 5.0), xlim = c(-1.5, 2.0), col = "blue", main = "Beauty and Evaluation for Males", xlab = "Beauty", ylab = "Evaluation")
fm_males <- lm(eval ~ beauty, data = males)
abline(fm_males)
```

@@

```{r}
#| echo: false
#| message: false
#| label: fig-females-beauty-eval
#| fig-cap: Beaty and Evaluation scores that increase for female instructors by 0.08762.
# females graph of beauty and evaluation 
females <- subset(TeachingRatings, gender == "female") |> select(eval, beauty, gender)
plot(eval ~ beauty, data = females, ylim = c(2.0, 5.0), xlim = c(-1.5, 2.0), col = "red", main = "Beauty and Evaluation For Females", xlab = "Beauty", ylab = "Evaluation")
fm_females <- lm(eval ~ beauty, data = females)
abline(fm_females)

@@

@fig-females-beauty-eval-with-males-abline shows the graph of the beauty and evaluation correlation for female instructors with the line of best fit for the male instructors. According to the models, for men the beauty affects their evaluations more than the female instructors. 
```{r}
#| echo: false
#| message: false
#| label: fig-females-beauty-eval-with-males-abline
#| fig-cap: Female instructors beauty and evaluation with the male and female line of best fit 
# females graph of beauty and evaluation with males abline added 
plot(eval ~ beauty, data = females, ylim = c(2.0, 5.0), xlim = c(-1.5, 2.0), main = "Beauty and Evaluation For Females With Males Abline", xlab = "Beauty", ylab = "Evaluation")
abline(fm_females, col = "red")
abline(fm_males, col = "blue")
legend("bottomright", legend = c("Males", "Females"), col = c("blue", "red"), lwd = 2)
```

@@

## evaluation score vs. beauty
plot(eval ~ beauty, data = TeachingRatings, main = "Beauty and Evaluation For All Instructors ", xlab = "Beauty", ylab = "Evaluation")
fm <- lm(eval ~ beauty, data = TeachingRatings)
abline(fm, col = "green")

>>>>

# 2.3 Data Analysis 
The data set from AER was gathered by installing the AER package [@AER] in the RStudio [@rstudio] console, and then the package contents were extracted to only include the TeachingRatings data set in the "01-download_data.R" file under "/scripts". The data was then analyzed using the R programming language [@rcoreteam2022]. Additionally, the programming packages dplyr [@wickham2023] as well as tidyverse [@wickham2019],[@jeppson2021] and [@testthat] were used to analyze the data, test and create graphs.

@@

# males graph of beauty and evaluation 
males <- TeachingRatings %>%
  filter(gender == "male") %>%
  select(eval, beauty, gender)

ggplot(males, aes(x = beauty, y = eval)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  xlim(-1.5, 2.0) +
  ylim(2.0, 5.0) +
  labs(title = "Beauty and Evaluation for Males",
       x = "Beauty", y = "Evaluation")
```

@@

```{r}
#| echo: false
#| message: false
#| label: fig-females-beauty-eval
#| fig-cap: Beaty and Evaluation scores that increase for female instructors by 0.08762.
# females graph of beauty and evaluation

females <- TeachingRatings %>%
  filter(gender == "female") %>%
  select(eval, beauty, gender)

ggplot(females, aes(x = beauty, y = eval)) +
  geom_point(color = "red") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  xlim(-1.5, 2.0) +
  ylim(2.0, 5.0) +
  labs(title = "Beauty and Evaluation for Females",
       x = "Beauty", y = "Evaluation")

@@

@fig-females-beauty-eval-with-males-abline shows the graph of the beauty and evaluation correlation for female instructors with the line of best fit for the male instructors. According to the models, for men the beauty affects their evaluations more than the female instructors. 
```{r}
#| echo: false
#| message: false
#| label: fig-females-beauty-eval-with-males-abline
#| fig-cap: Female instructors beauty and evaluation with the male(blue) and female(red) line of best fit 
# females graph of beauty and evaluation with males abline added 
males <- subset(TeachingRatings, gender == "male") |> select(eval, beauty, gender)
fm_males <- lm(eval ~ beauty, data = males)

females <- subset(TeachingRatings, gender == "female") |> select(eval, beauty, gender)
fm_females <- lm(eval ~ beauty, data = females)

females <- TeachingRatings %>%
  filter(gender == "female") %>%
  select(eval, beauty, gender)

ggplot(females, aes(x = beauty, y = eval, color = gender)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(-1.5, 2.0) +
  ylim(2.0, 5.0) +
  labs(title = "Beauty and Eval for Females With Female And Male Abline",
       x = "Beauty", y = "Evaluation") +
  geom_abline(slope = coef(fm_males)[2], intercept = coef(fm_males)[1], color = "blue") +
  geom_abline(slope = coef(fm_females)[2], intercept = coef(fm_females)[1], color = "red") +
  scale_color_manual(values = c("red", "red")) +
  theme(legend.position = "bottomright") +
  guides(color = guide_legend(title = "Gender")) 
```

@@

## evaluation score vs. beauty
library(ggplot2)

data("TeachingRatings", package = "AER")

ggplot(TeachingRatings, aes(x = beauty, y = eval)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "purple") +
  labs(title = "Beauty and Evaluation For All Instructors",
       x = "Beauty", y = "Evaluation")
